---
title: "k-means"
author: "Harika"
date: "4/25/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#libraries used

library(lubridate) # To work with dates and times
library(dplyr)     # To manipulate the datasets
library(tidyverse) # To design grammar,data structures
library(caret)     # To steamline the process for predicting models
library(factoextra)# To extract and visualize the results of data analyses
```{r}
library(lubridate)
library(dplyr)
library(tidyverse)
library(caret)
library(factoextra)
```

#import data
```{r}
raw_data <- read.csv("C:/Users/hpenj/OneDrive/Desktop/evg_AIP/inner_join_cop.csv")  #Importing the data from the local drive
data <- raw_data  #Saved as the raw_data to data name
```

#data cleaning
```{r}
#data <- data %>%  rename(customer_id = ï..customer_id)    #renaming the customer_id

data$order_date <- as.Date(data$order_purchase_timestamp,format = "%d-%m-%y")   #date format 

colSums(is.na(data))    #missing values of each columns

analysis_date <- lubridate::as_date('2018-10-18') #Analysis date for calculating recency 

#length(unique(data$customer_id))                      #unique customer_id 
#length(unique(data$customer_unique_id))               
#length(unique(data$order_date))


#removing any negative values in payment 
data_clean <- data %>% mutate(Amount = replace(payment_value, payment_value<=0, NA))
data_clean <- data_clean %>% drop_na()

#Total value of the order is product of installments and value
data_rfm <- data_clean %>% mutate(payment= payment_installments *payment_value)
```

# calculating rfm values
Recency : How recently purchase made by customers(Days of inactivity)
Frequency : How often customer makes a purchase (Number of purchases)
Monetary : How much money customer spends on purchase (Revenue generated)
```{r}
RFM <- data_rfm %>% group_by(customer_unique_id) %>% 
  summarise(recency=as.numeric(analysis_date-max(order_date)),
            frequency=n(), monetary = sum(payment)) 
```
Next step after calculating  the RFM values is to cluster the customers accordingly
Here we are usng K-means clustering,in order to proceed clustering analysis we need to know the distribution and also perform scaling.
#Transformation and scaling
```{r}
#Histogram of the RFM values
hist(RFM$recency)
hist(RFM$frequency)
hist(RFM$monetary)

#Log transformation 
RFM$recency.log <- log(RFM$recency)
RFM$frequency.log <- log(RFM$frequency)
RFM$monetary.log <- RFM$monetary + 0.1 # can't take log(0), so add a small value to remove zeros
RFM$monetary.log <- log(RFM$monetary.log)

#Scaling the data 
RFM$recency.z <- scale(RFM$recency.log, center=TRUE, scale=TRUE)
RFM$frequency.z <- scale(RFM$frequency.log, center=TRUE, scale=TRUE)
RFM$monetary.z <- scale(RFM$monetary.log, center=TRUE, scale=TRUE)
```

#plots for any outliers
```{r}
#ggplot for log transforamtion values
plot_log <- ggplot(RFM, aes(x = frequency.log, y = monetary.log))+ geom_point(aes(colour = recency.log))+ scale_colour_gradient(name="Log-transformed Recency")+xlab("Log-transformed Frequency")+ ylab("Log-transformed Monetary Value")
plot_log
#scatter<- scatter + geom_point(aes(colour = recency.log))
#scatter <- scatter + scale_colour_gradient(name="Log-transformed Recency")
#scatter<- scatter+ xlab("Log-transformed Frequency")
#scatter <- scatter + ylab("Log-transformed Monetary Value")
#scatter

#ggplot for scaled values
plot_z <- ggplot(RFM, aes(x = frequency.z, y = monetary.z))+ geom_point(aes(colour = recency.z))+ scale_colour_gradient(name=" Recency")+ xlab(" Frequency") + ylab("Monetary Value")
plot_z
#scatter1 <- scatter1 + geom_point(aes(colour = recency.z))

#scatter1 <- scatter1 + scale_colour_gradient(name="Z-scored Recency")
#scatter1 <- scatter1 + xlab("Z-scored Frequency")
#scatter1 <- scatter1 + ylab("Z-scored Monetary Value of Customer")
#scatter1
```

#remove outliers
```{r}

RFM <- RFM %>% filter(monetary.log >0 )
```

#wss and kmeans
Apply K-means algorithm
```{r}
set.seed(123)
RFM_z <- RFM[,c(8:10)] #Filtering required columnss for analysis
# function to compute total within-cluster sum of square 
wss <- function(k) {
  kmeans(RFM_z, k, nstart = 10 )$tot.withinss
}

# Compute and plot wss for k = 2 to k = 7
k.values <- 2:7

# extract wss for 2-7 clusters
wss_values <- map_dbl(k.values, wss)

plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE, 
       xlab="Number of clusters",
       ylab="Total within-clusters sum of squares(wss)")

model <- kmeans(RFM_z, centers = 4, iter.max = 500, nstart = 10) 
#model

RFM$Cluster <- as.factor(model$cluster)


KMeans_Results <- RFM %>%
  group_by(Cluster) %>%
  summarise(
            'Recency Mean' = round(mean(recency)),
            'Frequency Mean' = scales::comma(round(mean(frequency))),
            'Monetary Mean' = scales::dollar(round(mean(monetary))),
           
            )

DT::datatable((KMeans_Results),
              rownames = FALSE) # Display cluster means to identify their value to the business
```

#Exporting data to local drive
```{r}
data_output <- inner_join(RFM,data_rfm)
write.csv(data_output,"C:/Users/hpenj/OneDrive/Desktop/evg_AIP/data_output3.csv")


```


